Modal Abstractions for Virtualizing Memory Addresses

### 2023-03-24 programlogic.tex
+

 One possible approach is to abstract away the physical page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) from the virtual pointsto relation via replacing it with read-only abstract map entries -- a ghost (abstract) map entry per \textsf{L}$_4$\_\textsf{L}$_{1}$ mappings -- which can be used as a resource ($ \sumwalkabs\vaddr\qfrac\paddr$) whose ownership ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$) ,and as a token to access to the physical page-table-walk pointsto (\textsf{L}$_4$\_\textsf{L}$_{1}$) stored in an address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}). We use \textsf{Iris}'s authorative ghost-map logical machinery
\[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  (\mathcal{W}_{64} \; \rightarrow_{\textrm{fin}}\;  (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
to realize these reasoning principles ($\mathcal{A}\textsf{bsPTableWalk}$).
\begin{remark}[Exchanging Page-Table-Walk-Summarization Mappings (Tokens) with Physical Page-Table-Walk Pointstos ]
  $\mathcal{A}\textsf{PTableWalk}$'s fragmental-ownership represents an individual key-value pair (per-virtual-address) which can be handled as a \textit{look-up} resource for an individual virtual address.
  To build more intuition, we can think of the use-case \textit{unmapping-pages} as collecting the fragmental map pairs, e.g. $\sumwalkabs\vaddr\qfrac\paddr$, from collecting the assertions, removing them from the ghost map, and then updating the page tables.
  Critically, physical full ownership ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) of the page tables, as a \textit{singleton resource}, lives in the central invariant keeping physical-page-table pointstos per-virtual address translation.
  With this mechanism we can separate the virtual-to-physical translation from the physical memory involved, and consequently
  \begin{itemize}
  \item virtual pointsto assertions, whose definition is shown in Figure \ref{fig:virtualpointstosharing}, can refer to their physical memory assertions representing their page-address translation in the physical-page-tables ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) which reside in the address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) using individual key-value pair, i.e. instances of ($\sumwalkabs\vaddr\qfrac\paddr$)
  \item unlike the case in the strong virtual-pointsto relation in Figure \ref{fig:strongvirtualpointsto}, physical-page-table-walk pointsto assertions ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) for all virtual addresses are moved into the address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}), and resides with authorative ownership of $\mathcal{A}\textsf{PTableWalk}$ ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) which respects all entries of the abstract virtual-address translation mappings ($\sumwalkabs\vaddr\qfrac\paddr$).
    \end{itemize}
\end{remark}
\begin{figure*}
\[
\begin{array}{l}
    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vpage \stackrel{\triangle}{=} 
  \exists \entryo \ldotp 
  \sumwalkabs\vaddr\qfrac\paddr \ast 
   \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
\end{array}
\]
\caption{Virtual-Pointsto for Sharing Pages}
  \label{fig:virtualpointstosharing}
\end{figure*}
\subsection{Address-Space Abstraction}
\label{sec:aspacemodal}
Abstractions we introduce for virtual address spaces are based on the following aspects pictorially shown in Figure \ref{fig:logicaladdrspace}:
\begin{enumerate}

\item a global address space invariant (solid round square labelled as $\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) asserting the authorative ownership of the ghost map, $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$ \textit{represents} the full-ownership (solid-round-head arrow) of $\mathcal{A}\textsf{PTableWalk}$
\item per-virtual-pointsto relation ((cutted round square labelled as \textsf{Virtual PointsTo} in Figure \ref{fig:peraspaceinvariant})) asserting the fragmental ownership of the ghost map, $\sumwalkabs\vaddr\qfrac\paddr$ \textit{represents} the fragmental-ownership (cutted-diamond-head arrow) of $\mathcal{A}\textsf{PTableWalk}$
\item for each ghost page-table-walk mapping,  $\sumwalkabs\vaddr\qfrac\paddr$, there exists one shared persistent invariant (shown as double-ended solid arrow 1 to N)
  \[\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo)\]
  \item all virtual pointsto assertions happen to be true in a certain address space, i.e. the truth on address virtualization is indexed by CPUs page table register ($\crthree$ for x86-64), and we present modal reasoning principles for updating $\crthree$ (switching from the current address space to another)  

@@

>>>>

 One possible approach is to abstract away the physical page-table-walk pointsto ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$ in Figure \ref{fig:strongvirtualpointsto}) from the virtual pointsto relation via replacing it with read-only abstract map entries -- a ghost (abstract) map entry per \textsf{L}$_4$\_\textsf{L}$_{1}$ mappings -- which can be used as a resource ($ \sumwalkabs\vaddr\qfrac\paddr$) whose ownership ensures that the virtual address ($\vaddr$) is mapped to a page (referenced by $\paddr$) ,and as a token to access to the physical page-table-walk pointsto (\textsf{L}$_4$\_\textsf{L}$_{1}$) stored in an address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}). We use \textsf{Iris}'s authorative ghost-map logical machinery
\[\mathcal{A}\textsf{bsPTableWalk} \stackrel{\triangle}{=} \textsc{Auth} (\; \mathcal{W}_{64} \;\rightarrow_{\textrm{fin}} \;  (\mathcal{W}_{64} \; \rightarrow_{\textrm{fin}}\;  (\textsc{Frac }, \mathord{+}) \times (\textsc{Agree } \Loc,\mathord{=}) ))\]
to realize these reasoning principles ($\mathcal{A}\textsf{bsPTableWalk}$).
\begin{remark}[Exchanging Page-Table-Walk-Summarization Mappings (Tokens) with Physical Page-Table-Walk Pointstos ]
  $\mathcal{A}\textsf{PTableWalk}$'s fragmental-ownership represents an individual key-value pair (per-virtual-address) which can be handled as a \textit{look-up} resource for an individual virtual address.
  To build more intuition, we can think of the use-case \textit{unmapping-pages} as collecting the fragmental map pairs, e.g. $\sumwalkabs\vaddr\qfrac\paddr$, from collecting the assertions, removing them from the ghost map, and then updating the page tables.
  Critically, physical full ownership ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) of the page tables, as a \textit{singleton resource}, lives in the central invariant keeping physical-page-table pointstos per-virtual address translation.
  With this mechanism we can separate the virtual-to-physical translation from the physical memory involved, and consequently
  \begin{itemize}
  \item virtual pointsto assertions, whose definition is shown in Figure \ref{fig:virtualpointstosharing}, can refer to their physical memory assertions representing their page-address translation in the physical-page-tables ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) which reside in the address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) using individual key-value pair, i.e. instances of ($\sumwalkabs\vaddr\qfrac\paddr$)
  \item unlike the case in the strong virtual-pointsto relation in Figure \ref{fig:strongvirtualpointsto}, physical-page-table-walk pointsto assertions ($\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}$) for all virtual addresses are moved into the address-space invariant ($\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}), and resides with authorative ownership of $\mathcal{A}\textsf{PTableWalk}$ ($\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$) which respects all entries of the abstract virtual-address translation mappings ($\sumwalkabs\vaddr\qfrac\paddr$).
    \end{itemize}
\end{remark}
\begin{figure*}
\[
\begin{array}{l}
    \vaddr\mapsto_{\textsf{v}}\{\textsf{q}\}\vpage \stackrel{\triangle}{=} 
  \exists \entryo \ldotp 
  \sumwalkabs\vaddr\qfrac\paddr \ast 
   \nfpointsto{\mask\vaddr\ft\entryo}{\mask\vaddr\tw\entryo}\vpage\qfracone\naddr
\end{array}
\]
\caption{Virtual-Pointsto for Sharing Pages}
  \label{fig:virtualpointstosharing}
\end{figure*}
%\subsection{Address-Space Abstraction}
%\label{sec:aspacemodal}

Then, once we want to outline the distribution of the facts per-address, as we see it pictorially  in Figure \ref{fig:logicaladdrspace}, the outline is composed of the following aspects:
\begin{enumerate}
\item a global address space invariant (solid round square labelled as $\mathcal{I}$\textsf{ASpace} in Figure \ref{fig:peraspaceinvariant}) asserting the authorative ownership (i.e. update capability), $\ownGhost\gammaPred{\authfull{\ptableabswalk\ptablestore}}$, of the ghost map -- shown as solid-round-head arrow to $\mathcal{A}\textsf{PTableWalk}$ Figure \ref{fig:logicaladdrspace}
\item per-virtual-pointsto relation ((cutted round square labelled as \textsf{Virtual PointsTo} in Figure \ref{fig:peraspaceinvariant})) asserting the fragmental ownership of the ghost map, $\sumwalkabs\vaddr\qfrac\paddr$ \textit{represents} the fragmental-ownership (cutted-diamond-head arrow) of $\mathcal{A}\textsf{PTableWalk}$
\item for each ghost page-table-walk mapping,  $\sumwalkabs\vaddr\qfrac\paddr$, there exists one shared persistent invariant (shown as double-ended solid arrow 1 to N)
  \[\textsf{L}_{4}\_\textsf{L}_{1}\_\textsf{PointsTo}(\crthree,\entryf,\entrytr,\entrytw,\entryo)\]
\end{enumerate}

@@

\subsection{Address-Space Management}
\label{sec:aspacemanagement}
% \item abstractions shaped-around address-space switch where all virtual pointsto assertions happen to be true in a certain address space, i.e. the truth on address virtualization is indexed by CPUs page table register ($\crthree$ for x86-64), and we present modal reasoning principles for updating $\crthree$ (switching from the current address space to another) 
